---
title: "Figure Scratchbook"
author: "Kate Culhane"
output:
  html_document:
    theme: flatly
    highlight: tango
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE,
                      fig.width=5, fig.height=3)
```

**NOTE:** This is just a working space for ideas and static figures before I incorporate them into the Shiny app.

---


```{r packages and data}

# Packages
library(tidyverse)
library(janitor) # clean column names
library(kableExtra) # pretty tables
library(leaflet) # interactive maps
library(sf) # spatial functions
library(ISOcodes) # ISO country/province codes

# Raw personal eBird data
dat_personal <- read_csv("./data/MyEBirdData.csv") %>%
  clean_names() %>% 
  mutate(year = year(date))

# eBird taxonomy 
dat_taxonomy <- read_csv("./data/eBird_taxonomy_v2024.csv") %>% 
  clean_names() %>%
  separate(family, c("family", "family_common_name"), sep = "\\(|\\)") %>% 
  select(-taxon_concept_id, -report_as)

# ISO codes
iso_country <- ISO_3166_1 %>% 
  clean_names() %>% 
  mutate(country = if_else(is.na(common_name), name, common_name)) %>% 
  select("country_code"=alpha_2, country)
iso_province <- ISO_3166_2 %>% 
  rename(province_code = Code, province = Name, province_type = Type) %>% 
  mutate(country_code = str_sub(province_code, end=2)) %>% 
  select(-Parent)
dat_iso <- full_join(iso_country, iso_province, by = "country_code")

```

```{r data wrangling}

# Combine personal data, taxonomy, and ISO codes

dat_all <- left_join(dat_personal, dat_taxonomy,
                     by = c("scientific_name"="sci_name")) %>% 
  left_join(dat_iso,
            by = c("state_province"="province_code")) %>% 
  select(-primary_com_name, -country_code)

# Observation point data
dat_sf <- st_as_sf(dat_all, coords = c("longitude","latitude")) %>% 
  st_set_crs(4326) # set CRS to WGS84

```

## Notes on data

Personal data has one row per species per checklist.
Taxonomic data from [eBird](https://www.birds.cornell.edu/clementschecklist/introduction/updateindex/october-2024/2024-citation-checklist-downloads/?__hstc=60209138.90d3e80156bd0bf8cea9061b5c8b0677.1737839952332.1737839952332.1737839952332.1&__hssc=60209138.2.1737839952335&__hsfp=3572963934&_ga=2.47631268.2048720750.1737839952-547846681.1737839951&_gl=1*1r2sx3a*_gcl_au*MTk1NTg3NjU4OC4xNzM3ODM5OTUw*_ga*NTQ3ODQ2NjgxLjE3Mzc4Mzk5NTE.*_ga_QR4NVXZ8BM*MTczNzgzOTk1MC4xLjEuMTczNzgzOTk1Mi41OC4wLjA).

[Conditional function](https://forum.posit.co/t/conditional-pipelines/6076)

Data cleaning to do:
- Remove observations of upper-level taxa (e.g., Buteo sp.)
- Combine observations of subspecies

## Figure ideas

- Interactive map of observations
- Histogram of times checklists are submitted - but instead of end times have it include the entire span 
- Jason's county competition
- Word map of comments
- Diversity metrics
- Map through time gif

Toggle by taxonomic groups

## Top ten most frequently observed species

```{r top ten}

# Summary table

summary <- dat_all %>% 
  group_by(common_name, scientific_name) %>% 
  summarize(n = length(common_name)) %>% 
  ungroup() %>% 
  arrange(-n) %>% 
  top_n(10)

kable(summary[,1:3], col.names = c("Common name", "Latin binomial", "Observations")) %>% 
  kable_styling(bootstrap_options = c("striped", "condensed"),
                full_width = FALSE,
                position = 'left') %>% 
  column_spec(2, italic = TRUE) %>% 
  column_spec(1, width = '5cm')

```

## Interactive map

```{r interactive map}

# DROP-DOWN SELECTON OPTIONS

# Species group
list_sppgroup <- dat_all %>% 
  pull(species_group) %>% 
  unique()

# Species
list_spp <- dat_all %>% 
  filter(species_group=="Waterfowl") %>% # FILTER BY PREVIOUS
  pull(common_name) %>% 
  unique()

# Country
list_country <- dat_all %>% 
  pull(country) %>% 
  unique() %>% 
  str_sort()

# Province
list_province <- dat_all %>% 
  filter(country=="United States") %>% # FILTER BY PREVIOUS
  pull(province) %>%
  unique() %>% 
  str_sort()

# Year
list_year <- dat_all %>% 
  pull(year) %>% 
  unique() %>% 
  str_sort()

# INTERACTIVE MAP

dat_sf %>% 
  # Filter
  filter(species_group=="Waterfowl") %>%
  # filter(common_name=="Canada Goose") %>% 
  # Data wrangling
  group_by(submission_id, location, geometry) %>% 
  summarise() %>% 
  group_by(location, geometry) %>% 
  summarise(n = length(location)) %>% 
  arrange(-n) %>% 
  ungroup() %>% 
  mutate(loggedn = ifelse(n==1, 0.2/log(max(n),10), # change n=1 obs
                          log(n,10)/log(max(n),10))) %>% 
  arrange(-n) %>% 
  # Make map
  leaflet() %>% 
  addProviderTiles(providers$Esri.WorldGrayCanvas) %>%  
  addCircleMarkers(
    radius = ~loggedn*10+3,
    # color = ~pal(type),
    stroke = FALSE, fillOpacity = 0.5,
    label = ~location,
    popup = ~paste0(location,"<br>Checklists: ",n)
  )

```

Potentially consider only redrawing layers that change using [leafletProxy](https://rstudio.github.io/leaflet/articles/shiny.html) - could stop map from re-zooming every time a switch is toggled

